Combining neural networks and geostatistics for landslide hazard assessment of San Salvador Metropolitan Area, El Salvador \newline \newline Combinando redes neuronales y geoestad\'{i}stica para evaluaci\'{o}n de deslizamientos de tierra de el \'{a}rea Metropolitana de San Salvador, El Salvador

### 2015-06-05 paper.tex
04 June 2015 22:23

Throughout the year the country experiences a tropical climate with two seasons, a dry season (November to April) and wet season (May - October). The climate of El Salvador is generally warm. In the dry season there is very less rainfall but during rainy seasons heavy showers take place. El Salvador's interior regions remain dry throughout the year. Periodically El Salvador is affected by tropical storms and occasionally by hurricanes.

@@

 \begin{center}
%  \begin{figure}[H]
  \begin{figure}
   \centering
   \includegraphics[scale=0.75]{MASS_mapa_1}
   \caption{\small{Maps from El Salvador}}
   \label{fig:mass01}
  \end{figure}
 \end{center}

@@

\section{Landslide inventory and data sources of input variables}
\label{sec:landsinvet}
The Landslide inventory was developed by MARN and it was conducted in two stages. The first was carried out, based on fieldwork after the 2001 eartquakes in El Salvador. The last stage is complete with a visual inventory of places where there were landslides in visible satellite image IRSC (India) sensor of the same year.

from photographic analysis on the study area, using aerial photo from MARN which is presented in the Figure \ref{fig:img01}, where white areas represent regions of landslide occurrence which were processed using the software ILWIS.

@@

Once that the white areas were georeferenced, a percentage of 0.5\% of the total points georeferenced shows landslide ocurrence. In addition to the landslide information the following data sources of input variables was given by MARN: 
\begin{enumerate}
\item {\bf Geomorphology:} refers to landforms that result from lithospheric dynamics of geographic area.

\item {\bf Slope:} derived from digital elevation model MARN

\item {\bf Geology:} Description of the geology of the area from the map German Geological Mission (scale 1:100,000).

\item {\bf Rainfall intensity:} Maximum rainfall recorded in the geographical area.

\item {\bf Peak ground accelaration:} Maximum ground acceleration expressed in Gal for a return period of 500 years, this is the least that has detailed information. This information was obtained from RSIS II project.

\item {\bf Road distance:} Distance in kilometers to the nearest road.

\item {\bf Fault distance:} Distance in kilometers to the nearest fault. This information was obtained from German Geological Mission (scale 1:100,000).

@@

\begin{table}[H]
\caption{Summary of classification accuracy on the training, validation and test sets for neural network models  }
\label{annresults}
\centering
\begin{tabular}{ | c | c | c | c | }
\hline
\multicolumn{4}{| c |}{Percentage score} \\
\hline
Número de neuronas &    Train set    &   Validation set &  Test set \\
\hline
2  &  0.7011 & 0.7020 & 0.7124 \\
\hline
3 &  0.7206 & 0.7072 & 0.7401 \\
\hline 
4 &  0.7425 & 0.7197 & 0.7458 \\
\hline
5 &  0.7601 & 0.7604 & 0.7740 \\
\hline
6 &  0.7823 & 0.7704 & 0.7604 \\
\hline
7 &  0.7853  & 0.7704 &  0.7763 \\
\hline
8 &  0.7942 & 0.7865 & 0.7878  \\
\hline
9 & 0.8009 & 0.7871 & 0.7901  \\
\hline
10 & 0.8321 & 0.8132 & 0.8009  \\
\hline
11 & 0.8194 & 0.7792 & 0.7542 \\
\hline
12  &  0.8230 & 0.8006 & 0.8159 \\
\hline
13 & 0.8408 &  0.8006 & 0.8031 \\
\hline
14  &  0.8373 & 0.8017 & 0.8127 \\
\hline
15 &  0.8517 & 0.8210 & 0.8247 \\ 
\hline
16  &  0.8246 & 0.8017 & 0.8028 \\ 
\hline
17  &  0.8543 & 0.8283 & 0.8246 \\ 
\hline
18   &  0.8467 & 0.8022 & 0.8418 \\ 
\hline
19  &  0.8341  & 0.7991 & 0.8274 \\
\hline
\end{tabular}
\end{table}

@@

To generate the map of landslide hazards, the following steps were followed based on geostatistics methodology: Exploratory analysis, Variogram modelling and Spatial prediction using ordinary kriging and validation of this results.

In order to achieve the above steps a R scrit was developed which can be downloaded in the following address:

@@

As regards the estimation of the variogram, the spherical, gaussian and exponential models were fitted but the validation of these models on data that were not taken into account in the process of estimating the variogram showed that the best model was the exponential, since that it showed a $ R^{2} = 0.42 $ against $ R^{2} = 0.40 $ and $ R^{2} = 0.39 $ of the models spherical and gaussian respectively. With the exponential model were made spatial predictions on a grid of 1500 $\times$ 1500 and after that, the spatial predictions were imported to Geotiff raster format with the purpose of using this raster map in ArcGIS to generate the landslide hazard map for MASS.

@@

The results obtained in the geostatistics methodology showed that the spatial the data satisfies the conditions for applying kriging method. Final map of landslide probability of occurrence (see Fig \ref{fig:mass02} ) generated by kriging method is the result of multidisciplinary work which implied an intense geological analysis (i.e. MARN inventory of more than 4000 landslides) together with the use of sophisticated statistical techniques. This landslide hazard map is a user-friendly tool to evaluate landslide risk in MASS and can be useful for many purposes such as territorial planning, prevention and landslide mitigation and so on.

\begin{center}
%  \begin{figure}[H]
  \begin{figure}
   \centering
   \includegraphics[scale=0.75]{MASS_mapa_2}
   \caption{\small{Landslide hazard map MASS}}
   \label{fig:mass02}
  \end{figure}
 \end{center}

High probability of landslide occurrence in Picacho slope implies very high risk due the proximity of urban areas and due the large volume of mobilized materials.(\cite{major2004}) identified in San Salvador volcano that lahars as voluminous as 2 million of m3 can affect large zones of urban areas. Southern areas of MASS also implies a high risk because large landslides, similar to Las Colinas landslide which involved a total volume of around 150,000 m3 (\cite{evans}), can occur in northern slope of Balsamo range and San Jacinto Hill affecting very crowded urban areas.

The present hazard landslide map is consistent with former landslide hazard maps carried out in the region using approaches conceptually different. (\cite{fernan2008}) presented a landslide hazard GIS map of MASS generated through bivariate statistical method considering several conditioning factors showing similar hazard areas. The only differences are located mainly in the Nejapa Hill, San Jacinto Hill where (\cite{fernan2008}) identifies a slightly higher hazard and the area between San Jacinto Hill and Ilopango Lake where the former study presents a hazard slightly lower. As landslides are an erosional phenomenon, the highest probability areas of landslide occurrence coincides with the areas with highest hazard of erosion in the erosion hazard map of MASS proposed by (\cite{chavez2014a}). There are similarities with susceptibility hazard maps at national scale generated through logistic regression method (\cite{garcia2008}), through artificial neural network method (\cite{garcia2010}) and Mora-Vahrson method (\cite{snet2004}). However the high landslide areas are more extended in these national maps than the present study. These differences could be explained by the working scale, our map is more detailed because is at regional scale. Other possible explanation can be related to the data, present work was built using a more extended landslide catalog and more updated data (e.g. seismic hazard from \cite{beni2012}) than former works.

@@

\section{Acknowledgment}   
The authors are indebted with MARN who provided the data and GIS software such as ILWIS and ARCGIS.

@@

\bibitem{weber}
Weber H.S., Wiesemann G., Lorenz  W., Schmidt-Thom\'{e} M.
\newblock Mapa geol\'{o}gico de Republica de El Salvador
\newblock escala 1 : 100 000 (6 sheets).
\newblock \emph{ Bundesanstalt f\"ur Geowissenschaften und Rohstoffe, Hannover, Germany. }

\bibitem{bommer}
Bommer J.J., Rodriguez C.E.
\newblock Earthquake-induced landslides in Central Am\'{e}rica
\newblock \emph{Engineering Geology}.
(63):\penalty0 1889-220, 2002

@@

>>>>

Throughout the year the country experiences a tropical climate with two seasons, a dry season (November to April) and wet season (May - October). The climate of El Salvador is generally warm. In the dry season there is very less rainfall but during rainy seasons heavy showers take place. Periodically El Salvador is affected by tropical storms and occasionally by hurricanes.

@@

 \begin{center}
%  \begin{figure}[H]
  \begin{figure}
   \centering
   \includegraphics[scale=0.70]{MASS_mapa_1}
   \caption{\small{Location of El Salvador in Central America, map of El Salvador with location of the Area Metropolitana de San Salvador (MASS) and topography of MASS.}}
   \label{fig:mass01}
  \end{figure}
 \end{center}

@@

\section{Landslide inventory and data sources of input variables}
\label{sec:landsinvet}
The Landslide inventory was developed by MARN and it was conducted in two stages. The first was carried out, based on fieldwork after the 2001 eartquakes in El Salvador. The another stage was completed with a visual inventory of places where there were landslides which were visible in a image of satellite image from IRS-C sensor (Indian) in the same year.

%from photographic analysis on the study area, using aerial photo from MARN which is presented in the Figure \ref{fig:img01}, where white areas represent regions of landslide occurrence which were processed using the software ILWIS.

@@

Once that the landslide inventory was finished, a percentage of 0.5\% of the total points georeferenced shows landslide ocurrence. In addition to the landslide information the following data sources of input variables was provided by MARN:

\begin{enumerate}
\item {\bf Geomorphology:} refers to landforms that result from lithospheric dynamics of geographic area. The physical properties of slope-forming materials, such as strength and permeability, are related to the lithology factor, which therefore should affect the likelihood of slope failure (\cite{dai2002}). This variables was obtained in accordance with the project of national land use plan in El Salvador.

\item {\bf Slope:} Slope gradient is an important component and a
preparatory cause of landsliding (\cite{garcia2008}). It was calculated from digital elevation model (DEM) provided by MARN with a resolution of 10 mts.

\item {\bf Geology:} Description of the geology of the area from the map German Geological Mission (scale 1:100,000) (\cite{weber}).

\item {\bf Rainfall intensity:} From the precipitation database compiled by the
SNET's weather stations for the period 1970–2001, maximum precipitations were obtained, and then was applied interpolation kriging to them.

\item {\bf Peak ground accelaration:} Maximum ground acceleration expressed in Gal for a return period of 500 years. This data was obtained from the evaluation of seismic Hazard in Central America carried out under the framework of RSIS II project (\cite{beni2012}).

\item {\bf Road distance:} Distance in kilometers to the nearest road. Road map was obtained from digitalized national cartography from Centro Nacional de Registros de la República de El Salvador (CNR).

\item {\bf Fault distance:} Distance in kilometers to the nearest fault. This information was obtained from according digitalized geological map of El Salvador (scale 1:100,000) (\cite{weber}).

@@

\begin{table}[H]
\caption{Summary of classification accuracy on the training, validation and test sets for neural network models  }
\label{annresults}
\centering
\begin{tabular}{ | c | c | c | c | }
\hline
\multicolumn{4}{| c |}{Percentage score} \\
\hline
Number of neurons &    Train set    &   Validation set &  Test set \\
\hline
2  &  0.7011 & 0.7020 & 0.7124 \\
\hline
3 &  0.7206 & 0.7072 & 0.7401 \\
\hline 
4 &  0.7425 & 0.7197 & 0.7458 \\
\hline
5 &  0.7601 & 0.7604 & 0.7740 \\
\hline
6 &  0.7823 & 0.7704 & 0.7604 \\
\hline
7 &  0.7853  & 0.7704 &  0.7763 \\
\hline
8 &  0.7942 & 0.7865 & 0.7878  \\
\hline
9 & 0.8009 & 0.7871 & 0.7901  \\
\hline
10 & 0.8321 & 0.8132 & 0.8009  \\
\hline
11 & 0.8194 & 0.7792 & 0.7542 \\
\hline
12  &  0.8230 & 0.8006 & 0.8159 \\
\hline
13 & 0.8408 &  0.8006 & 0.8031 \\
\hline
14  &  0.8373 & 0.8017 & 0.8127 \\
\hline
15 &  0.8517 & 0.8210 & 0.8247 \\ 
\hline
16  &  0.8246 & 0.8017 & 0.8028 \\ 
\hline
17  &  0.8543 & 0.8283 & 0.8246 \\ 
\hline
18   &  0.8467 & 0.8022 & 0.8418 \\ 
\hline
19  &  0.8341  & 0.7991 & 0.8274 \\
\hline
\end{tabular}
\end{table}

@@

To generate the map of landslide hazards, the following steps were followed based on geostatistics methodology: Exploratory analysis, Variogram modelling and Spatial prediction using ordinary kriging, validation of this results and .

In order to achieve the above steps a R script was developed which can be downloaded in the following address:

@@

As regards the estimation of the variogram, the spherical, gaussian and exponential models were fitted but the validation of these models on data that were not taken into account in the process of estimating the variogram showed that the best model was the exponential, since that it showed a $ R^{2} = 0.42 $ against $ R^{2} = 0.40 $ and $ R^{2} = 0.39 $ of the models spherical and gaussian respectively. With the exponential model were made spatial predictions on a grid of 1500 $\times$ 1500 and after that, the spatial predictions were exported to Geotiff raster format. Finally Geographic Information System (ArcGIS\textregistered \  of ESRI) was used to design the final map (Figure \ref{fig:mass02}). This map (Fig. \ref{fig:mass02}) represents the landslide probability of occurrences in MASS. According to Fig. \ref{fig:mass02} higher probability of landslide occurrence, reaching probability of occurrence values around 0.8, are located in the northern slopes of Balsamo Range (western side of MASS), in the eastern slope of San Salvador volcano and Picacho,  along the extension of Mariona Hills  and in the southeastern side, between Ilopango lake and San Jacinto Hill. Areas with lower probability but still important, are located in southern and western side of Nejapa Hill, in north-east of the study area, in northern banks of Ilopango lake, in northeast slope of San Salvador Volcano, in the surrounding areas of San Jacinto Hill and, in general in southern side of MASS.

@@

The results obtained in the geostatistics methodology showed that the spatial the data satisfies the conditions for applying kriging method. Final map of landslide probability of occurrence (see Fig \ref{fig:mass02} ) generated by kriging method is the result of multidisciplinary work which implied an intense geological analysis (i.e. MARN inventory of more than 4000 landslides) together with the use of sophisticated statistical techniques. This landslide hazard map is a user-friendly tool to evaluate landslide risk in MASS and can be useful for many purposes such as territorial planning, prevention and landslide mitigation and so on. 
\begin{center}
%  \begin{figure}[H]
  \begin{figure}
   \centering
   \includegraphics[scale=0.70]{MASS_mapa_2}
   \caption{\small{Landslide hazard map MASS where is shown the landslide probability of occurrence.}}
   \label{fig:mass02}
  \end{figure}
 \end{center}
High probability of landslide occurrence in Picacho slope implies very high risk due the proximity of urban areas and due the large volume of mobilized materials.(\cite{major2004}) identified in San Salvador volcano that lahars as voluminous as 2 million of m3 can affect large zones of urban areas. Southern areas of MASS also implies a high risk because large landslides, similar to Las Colinas landslide which involved a total volume of around 150,000 m3 (\cite{evans}), can occur in northern slope of Balsamo range and San Jacinto Hill affecting very crowded urban areas.

The present hazard landslide map is consistent with former landslide hazard maps carried out in the region using approaches conceptually different. (\cite{fernan2008}) presented a landslide hazard GIS map of MASS generated through bivariate statistical method considering several conditioning factors showing similar hazard areas. The only differences are located mainly in the Nejapa Hill, San Jacinto Hill where (\cite{fernan2008}) identifies a slightly higher hazard and the area between San Jacinto Hill and Ilopango Lake where the former study presents a hazard slightly lower. As landslides are an erosional phenomenon, the highest probability areas of landslide occurrence coincides with the areas with highest hazard of erosion in the erosion hazard map of MASS proposed by (\cite{chavez2014a}). There are similarities with susceptibility hazard maps at national scale generated through logistic regression method (\cite{garcia2008}), through artificial neural network method (\cite{garcia2010}) and through Mora-Vahrson method (\cite{snet2004}). However the high landslide areas are more extended in these national maps than the present study. These differences could be explained by the working scale, our map is more detailed because is at regional scale. Other possible explanation can be related to the data, present work was built using a more extended landslide catalog and more updated data (e.g. seismic hazard evaluation from \cite{beni2012}) than former works.

@@

\section{Acknowledgment}   
The authors are indebted with MARN who provided the data and Geogrphic Information System software such as ILWIS and ArcGIS\textregistered \  of ESRI.

@@

\bibitem{weber}
Weber H.S., Wiesemann G., Lorenz  W., Schmidt-Thom\'{e} M.
\newblock Mapa geol\'{o}gico de Republica de El Salvador
\newblock escala 1 : 100 000 (6 sheets).
\newblock \emph{ Bundesanstalt f\"ur Geowissenschaften und Rohstoffe, Hannover, Germany.
}
\penalty0 (1978).

\bibitem{bommer}
Bommer J.J., Rodriguez C.E.
\newblock Earthquake-induced landslides in Central Am\'{e}rica
\newblock \emph{Engineering Geology}.
(63):\penalty0 1889-220, (2002).

@@

\bibitem{dai2002}
Dai, F. C. and Lee, C. F.
\newblock Landslide characteristics and slope instability modeling using GIS, Lantau Island, Hong Kong.
\newblock \emph{Geomorphology}
\penalty0 42, 213–228, 2002.
